Understanding and Verifying the Shape of Stencil Computations

### 2017-03-27 oopsla17/lattice-model.tex
Add lemma proof + stylistic fixes
Modify region space to region directly

% Essential macros
\newcommand{\zinf}{\textnormal{$\mathbb{Z}_\infty$}}
\newcommand{\interp}[1]{\llbracket {#1} \rrbracket}
\newcommand{\interv}[3]{\textnormal{$\lbrack#1,#2\rbrack^{#3}$}}
\newcommand{\vecgen}[3]{\textnormal{$\textit{vec-gen}_{#1}(#2,#3)$}}
\newcommand{\textcap}[1]{\texttt{\textit{#1}}}

% Stencil macro
%
% Optional argument: s when we want the short form, empty otherwise.
%
% Mandatory arguments:
% 1st: c/b/f/p for centered, backward, forward, pointed
% 2st: dimension
% 3st: depth
% 4st: if pointed/nonpointed needs to be specified leave it blank if omitted.
%
% Example: \stencil[s]{f}{1}{2}{} = fwd(dim=1,depth=2)
% Example: \stencil{b}{1}{2}{pointed} = backward(dim=1,depth=2,pointed)
\newcommand{\stencil}[5][]{%
  \ifthenelse{\equal{#2}{f}}{%
    \ifthenelse{\equal{#1}{s}}{%
      \texttt{fwd}%
    }{\ifthenelse{\isempty{#1}}{%
        \texttt{forward}%
      }{INVALID OPTIONAL}%
    }%
  }{\ifthenelse{\equal{#2}{b}}{%
      \ifthenelse{\equal{#1}{s}}{%
          \texttt{bwd}%
      }{\ifthenelse{\isempty{#1}}{%
          \texttt{backward}%
        }{INVALID OPTIONAL}%
      }%
  }{\ifthenelse{\equal{#2}{c}}{%
      \ifthenelse{\equal{#1}{s}}{%
          \texttt{ctd}%
      }{\ifthenelse{\isempty{#1}}{%
          \texttt{centered}%
        }{INVALID OPTIONAL}%
      }%
  }{\ifthenelse{\equal{#2}{p}}{%
      \ifthenelse{\equal{#1}{s}}{%
          \texttt{ptd}%
      }{\ifthenelse{\isempty{#1}}{%
          \texttt{pointed}%
        }{INVALID OPTIONAL}%
      }%
  }{\ifthenelse{\isempty{#1}}{%
      \textcap{#2}%
    }{NO OPTIONAL FOR CUSTOM}%
  }}}}%
  \texttt{(dim=\textnormal{#3}%
    \ifthenelse{\isempty{#4}}{}{,depth=\textnormal{#4}}%
    \ifthenelse{\isempty{#5}}{}{,\textnormal{#5}}%
  )}%
}

@@

\unused{
\begin{lem}{}\label{lem:vector-intersect}
  Intersection distributes over offsets vectors.
%
  \begin{equation*}
    (\textit{vint}_1, \dots, \textit{vint}_N) \;\cap\;
      (\textit{vint'}_1, \dots, \textit{vint'}_N)
      =
    (\textit{vint}_1 \;\cap\; \textit{vint'}_1, \dots,
     \textit{vint}_N \;\cap\; \textit{vint'}_N)
  \end{equation*}
\end{lem}
%
\begin{proof}
  TODO. Trivial. Do if needed.
\end{proof}
}

\begin{lem}{}\label{lem:vector-union}
  Let $v$ and $v'$ be two $N$ dimensional offsets vectors such that $v_i$ is
  equal to $v'_i$ for all $1 \leq i \leq N$ except $k$, then we have the
  following:
%
  \begin{equation*}
    (v_1,\dotsc, v_k,\dotsc, v_N) \; \cup \; (v'_1,\dotsc, v'_k, \dotsc, v_N)
    =
    (v_1,\dotsc, v_k \cup v'_k,\dotsc, v_N)
  \end{equation*}
\end{lem}
%
\begin{proof}
  \begin{align*}
    v \; \cup \; v' &
    = \{(x_1,\dotsc,x_n) \mid
          \bigwedge_{1 \leq i \leq N } x_i \in v_i \}
      \;\cup\;
      \{(x_1,\dotsc,x_n) \mid
          \bigwedge_{1 \leq i \leq N } x_i \in v'_i \} \\
    & = \{(x_1,\dotsc,x_n) \mid
          \bigwedge_{\substack{1 \leq i \leq N \\ i \neq k}} x_i \in v_i
              \wedge x_k \in v_k \vee
          \bigwedge_{\substack{1 \leq i \leq N \\ i \neq k}} x_i \in v'_i
            \wedge x_k \in v'_k
        \} \\
    & = \{(x_1,\dotsc,x_n) \mid
          \bigwedge_{\substack{1 \leq i \leq N \\ i \neq k}} x_i \in v_i
          \wedge x_k \in v_k \cup v'_k
        \} \\
    & = (v_1,\dotsc, v_k \cup v'_k,\dotsc, v_N)
  \end{align*}
\end{proof}

@@

\unused{
  \begin{lem}{}
    Intersection distributes $N$ dimensional vector intervals:
  %
    \begin{gather*}
      (\interv{a_1}{b_1}{c_1}, \dots, \interv{a_N}{b_N}{c_N}) \; \cap \;
        (\interv{a'_1}{b'_1}{c'_1}, \dots, \interv{a'_N}{b'_N}{c'_N}) \\
      = \\
      (\interv{\max \{a_1,a'_1\}}{\min \{b_1,b'_1\}}{\neg c_1 \vee \neg c'_1},
        \dots, \interv{\max \{a_N,a'_N\}}{\min \{b_N,b'_N\}}{\neg c_N \vee \neg c'_N})
    \end{gather*}
  \end{lem}
  %
  \begin{proof}
    This immediately follows from \Cref{lem:vector-intersect} and
    \Cref{lem:zinf-identities}.
  \end{proof}
}

\begin{defn}{(\emph{Region space})}
  Region space defines the set \textit{Region} in $N$ dimensions inductively as
  follows:
%
  \begin{enumerate}
    \item If $R$ is in $\textit{Interval}^N$, then $R$ is in \textit{Region}.
    \item If $R$ and $S$ are in \textit{Region}, then so are $R \cap S$ and $R
      \cup S$.
  \end{enumerate}
\end{defn}

@@

        \hspace{2em}&\hspace{-2em} \interp{

@@

        &= \interp{

@@

      To show that the intersection is \textit{ks}, we use
      \Cref{lem:vector-intersect} which reduces the problem to showing that the
      intersection of $\textit{int}_k$ and $\textit{int}_l$ is the former since
      at every point except $n$ the intervals agree, hence the intersection
      remains the same. By assumption we have $k \leq l$, so the maximum of the
      lower bounds is $-k$ and the minimum of lower bounds is $k$ this is same
      as interval $\textit{int}_k$. Note that pointedness is the same for both
      intervals and depths are positive integers, so the pointedness is
      preserved under intersection. This shows the intersection of \textit{ls}
      and \textit{ks} is indeed \textit{ks} and concludes the proof.
  \end{description}
\end{proof}

>>>>

% Essential macros
\newcommand{\zinf}{\textnormal{$\mathbb{Z}_\infty$}}
\newcommand{\interp}[1]{\llbracket{} {#1} \rrbracket{}}
\newcommand{\interv}[3]{\textnormal{$\lbrack#1,#2\rbrack^{#3}$}}
\newcommand{\vecgen}[3]{\textnormal{$\textit{vec-gen}_{#1}(#2,#3)$}}
\newcommand{\textcap}[1]{\texttt{\textit{#1}}}

% Stencil macro
%
% Optional argument: s when we want the short form, empty otherwise.
%
% Mandatory arguments:
% 1st: c/b/f/p for centered, backward, forward, pointed
% 2st: dimension
% 3st: depth
% 4st: if pointed/nonpointed needs to be specified leave it blank if omitted.
%
% Example: \stencil[s]{f}{1}{2}{} = fwd(dim=1,depth=2)
% Example: \stencil{b}{1}{2}{pointed} = backward(dim=1,depth=2,pointed)
\newcommand{\stencil}[5][]{%
  \ifthenelse{\equal{#2}{f}}{%
    \ifthenelse{\equal{#1}{s}}{%
      \texttt{fwd}%
    }{\ifthenelse{\isempty{#1}}{%
        \texttt{forward}%
      }{INVALID OPTIONAL}%
    }%
  }{\ifthenelse{\equal{#2}{b}}{%
      \ifthenelse{\equal{#1}{s}}{%
          \texttt{bwd}%
      }{\ifthenelse{\isempty{#1}}{%
          \texttt{backward}%
        }{INVALID OPTIONAL}%
      }%
  }{\ifthenelse{\equal{#2}{c}}{%
      \ifthenelse{\equal{#1}{s}}{%
          \texttt{ctd}%
      }{\ifthenelse{\isempty{#1}}{%
          \texttt{centered}%
        }{INVALID OPTIONAL}%
      }%
  }{\ifthenelse{\equal{#2}{p}}{%
      \ifthenelse{\equal{#1}{s}}{%
          \texttt{ptd}%
      }{\ifthenelse{\isempty{#1}}{%
          \texttt{pointed}%
        }{INVALID OPTIONAL}%
      }%
  }{\ifthenelse{\isempty{#1}}{%
      \textcap{#2}%
    }{NO OPTIONAL FOR CUSTOM}%
  }}}}%
  \texttt{(dim=\textnormal{#3}%
    \ifthenelse{\isempty{#4}}{}{,depth=\textnormal{#4}}%
    \ifthenelse{\isempty{#5}}{}{,\textnormal{#5}})}%
}

@@

\begin{lem}{}\label{lem:vector-intersect}
  Intersection distributes over offsets vectors.
%
  \begin{equation*}
    (\textit{vint}_1, \dots, \textit{vint}_N) \;\cap\;
      (\textit{vint'}_1, \dots, \textit{vint'}_N)
      =
    (\textit{vint}_1 \;\cap\; \textit{vint'}_1, \dots,
     \textit{vint}_N \;\cap\; \textit{vint'}_N)
  \end{equation*}
\end{lem}
%
\begin{proof}
  \begin{align*}
    v \; \cap \; v' &
    = \{x \mid \bigwedge_{1 \leq i \leq N } x_i \in v_i \}
      \;\cap\;
      \{x \mid \bigwedge_{1 \leq i \leq N } x_i \in v'_i \} \\
    & = \{x \mid \bigwedge_{1 \leq i \leq N } (x_i \in v_i \wedge x_i \in v'_i) \} \\
    & = (v_1 \cap v'_1,\dotsc,v_N \cap v'_N)
  \end{align*}
\end{proof}

\begin{lem}{}\label{lem:vector-union}
  Let $v$ and $v'$ be two $N$ dimensional offsets vectors such that $v_i$ is
  equal to $v'_i$ for all $1 \leq i \leq N$ except $k$, then we have the
  following:
%
  \begin{equation*}
    (v_1,\dotsc, v_k,\dotsc, v_N) \; \cup \; (v'_1,\dotsc, v'_k, \dotsc, v_N)
    =
    (v_1,\dotsc, v_k \cup v'_k,\dotsc, v_N)
  \end{equation*}
\end{lem}
%
\begin{proof}
  \begin{align*}
    v \; \cup \; v' &
    = \{x \mid
          \bigwedge_{1 \leq i \leq N } x_i \in v_i \}
      \;\cup\;
      \{x \mid
          \bigwedge_{1 \leq i \leq N } x_i \in v'_i \} \\
    & = \{x \mid
          \bigwedge_{\substack{1 \leq i \leq N \\ i \neq k}} x_i \in v_i
              \wedge x_k \in v_k \vee
          \bigwedge_{\substack{1 \leq i \leq N \\ i \neq k}} x_i \in v'_i
            \wedge x_k \in v'_k
        \} \\
    & = \{x \mid
          \bigwedge_{\substack{1 \leq i \leq N \\ i \neq k}} x_i \in v_i
          \wedge x_k \in v_k \cup v'_k
        \} \\
    & = (v_1,\dotsc, v_k \cup v'_k,\dotsc, v_N)
  \end{align*}
\end{proof}

@@

\begin{defn}{(\emph{Region})}
  \textit{Region} is the set of $N$ dimensional vector intervals defined as the
  smallest set satisfying the following:
%
  \begin{enumerate}
    \item If $R$ is in $\textit{Interval}^N$, then $R$ is in \textit{Region}.
    \item If $R$ and $S$ are in \textit{Region}, then so are $R \cap S$ and $R
      \cup S$.
  \end{enumerate}
\end{defn}

@@

        \hspace{2em}&\hspace{-2em} \interp{%

@@

        &= \interp{%

@@

      To show that the intersection is \textit{ks}, we use
      \Cref{lem:vector-intersect} which reduces the problem to showing that the
      intersection of $\textit{int}_k$ and $\textit{int}_l$ is the former since
      at every point except $n$ the intervals agree, hence the intersection
      remains the same. Then we use \Cref{lem:zinf-identities} to calculate the
      intersection. By assumption we have $k \leq l$, so the maximum of the
      lower bounds is $-k$ and the minimum of lower bounds is $k$ this is same
      as interval $\textit{int}_k$. Note that pointedness is the same for both
      intervals and depths are positive integers, so the pointedness is
      preserved under intersection. This shows the intersection of \textit{ls}
      and \textit{ks} is indeed \textit{ks} and concludes the proof.
  \end{description}
\end{proof}
